---
title: "Homework_5"
author: "Ramazan Yarar"
date: "1/3/2019"
output: html_document
---


# Step 1
 
```{r message=FALSE,error=FALSE}
library(cluster)
library(MASS)
library(clusterGeneration)
library(factoextra)
library(rattle)
library(pdist)
library(data.table)
library(caret)
library(glmnet)

data_musk<-read.table("~/Documents/Dersler/Fall-2018/IE-582/Musk1.csv",header = FALSE, sep = ",")

data_musk<-cbind(data_musk[,1:2],scale(data_musk[,3:168]))

```

* Clustering


```{r cars}



kmedoids.musk<-pam(data_musk[,3:168],k=5)


# find best cluster count with elbow and silhouette methods



fviz_nbclust(data_musk[,3:168], pam, method = "wss") 

fviz_nbclust(data_musk[,3:168], hcut, method = "silhouette",dist(data_musk[,3:168], method = "euclidean"),
             hc_method = "ward.D")


```



 The elbow andsilhouette methods shows that the optimal number of cluster is 3. But I prefer 5 and 10 cluster to increase features to classification.  




```{r pressure, echo=FALSE}

# 5 cluster

# hierchical clustering

hclust.musk<- hcut(data_musk[,3:168], hc_method = "ward.D",hc_metric = "euclidean",k=5)

hclust.muskm<- hcut(data_musk[,3:168], hc_method = "ward.D",hc_metric = "minkowski",k=5)


hclust.musk_centreoids<-centers.hclust(data_musk[,3:168],hclust.musk, nclust=5, use.median=TRUE)

hclust.muskm_centreoids<-centers.hclust(data_musk[,3:168],hclust.muskm, nclust=5, use.median=TRUE)

#distance to centroids

dist_hclust_eucl<-pdist(data_musk[,3:168],hclust.musk_centreoids)@dist

dist_hclust_min<-pdist(data_musk[,3:168],hclust.muskm_centreoids)@dist


data_hclust_eucl<-cbind(data_musk[,1:2],data.table(matrix(dist_hclust_eucl,nrow = 476,ncol = 5)))


data_hclust_min<-cbind(data_musk[,1:2],data.table(matrix(dist_hclust_min,nrow = 476,ncol = 5)))



# k medoids

kmedoids.musk<-pam(data_musk[,3:168],k=5)

kmedoids_centroids<-kmedoids.musk$medoids

#distance to centroids

dist_kmedoids<-pdist(data_musk[,3:168],kmedoids_centroids)@dist

data_kmedoids<-cbind(data_musk[,1:2],data.table(matrix(dist_kmedoids,nrow = 476,ncol = 5)))

data_keybag<-data.table(unique(data_musk[,1:2]))




```



* Step 3


```{r}



data_kmedoids_ag<-data.table(aggregate(data_kmedoids[,3:7], list(data_kmedoids[,2]), mean))

data_hclust_eucl_ag<-data.table(aggregate(data_hclust_eucl[,3:7], list(data_hclust_eucl[,2]), mean))

data_hclust_min_ag<-data.table(aggregate(data_hclust_min[,3:7], list(data_hclust_min[,2]), mean))


data_kmedoids <- merge(data_kmedoids_ag,data_keybag,by.x = "Group.1", by.y = "V2",all.x =TRUE)

data_hclust_eucl <- merge(data_hclust_eucl_ag,data_keybag,by.x = "Group.1", by.y = "V2",all.x =TRUE)

data_hclust_min <- merge(data_hclust_min_ag,data_keybag,by.x = "Group.1", by.y = "V2",all.x =TRUE)



```





```{r}

summary=data.table(algorithms=character(), Auc=numeric(),FeatureCount=numeric())


x_medoids<- as.matrix(data_kmedoids[,2:6])
y_medoids<- data_kmedoids$V1.y

modelprakmedoids <- cv.glmnet(x_medoids, y_medoids, alpha = 1,nfolds=7,type.measure='auc',family="binomial")

summary<-rbind(summary,list("kmedoids",max(modelprakmedoids$cvm),5))

#max(modelprakmedoids$cvm)

x_hclust_eucl<-as.matrix(data_hclust_eucl[,2:6])
y_hclust_eucl<- data_hclust_eucl$V1.y

modelprahclust_eucl <- cv.glmnet(x_hclust_eucl, y_hclust_eucl, alpha = 1,nfolds=7,type.measure='auc',family="binomial")


summary<-rbind(summary,list("hclust_euclidean",max(modelprahclust_eucl$cvm),5))
#max(modelprahclust_eucl$cvm)


x_hclust_minl<-as.matrix(data_hclust_min[,2:6])
y_hclust_minl<- data_hclust_min$V1.y

modelprahclust_minl <- cv.glmnet(x_hclust_minl, y_hclust_minl, alpha = 1,nfolds=7,type.measure='auc',family="binomial")

summary<-rbind(summary,list("hclust_minkowski",max(modelprahclust_minl$cvm),5))




```



* 10 cluster

```{r}

hclust.musk10<- hcut(data_musk[,3:168], hc_method = "ward.D",hc_metric = "euclidean",k=10)

hclust.muskm10<- hcut(data_musk[,3:168], hc_method = "ward.D",hc_metric = "minkowski",k=10)


hclust.musk_centreoids10<-centers.hclust(data_musk[,3:168],hclust.musk, nclust=10, use.median=TRUE)

hclust.muskm_centreoids10<-centers.hclust(data_musk[,3:168],hclust.muskm, nclust=10, use.median=TRUE)

dist_hclust_eucl10<-pdist(data_musk[,3:168],hclust.musk_centreoids10)@dist

dist_hclust_min10<-pdist(data_musk[,3:168],hclust.muskm_centreoids10)@dist


data_hclust_eucl10<-cbind(data_musk[,1:2],data.table(matrix(dist_hclust_eucl10,nrow = 476,ncol = 10)))


data_hclust_min10<-cbind(data_musk[,1:2],data.table(matrix(dist_hclust_min10,nrow = 476,ncol = 10)))



# k medoids

kmedoids.musk10<-pam(data_musk[,3:168],k=10)

kmedoids_centroids10<-kmedoids.musk10$medoids


dist_kmedoids10<-pdist(data_musk[,3:168],kmedoids_centroids10)@dist

data_kmedoids10<-cbind(data_musk[,1:2],data.table(matrix(dist_kmedoids10,nrow = 476,ncol = 10)))

data_keybag<-data.table(unique(data_musk[,1:2]))



```


##




```{r}

data_kmedoids_ag10<-data.table(aggregate(data_kmedoids10[,3:12], list(data_kmedoids10[,2]), mean))

data_hclust_eucl_ag10<-data.table(aggregate(data_hclust_eucl10[,3:12], list(data_hclust_eucl10[,2]), mean))

data_hclust_min_ag10<-data.table(aggregate(data_hclust_min10[,3:12], list(data_hclust_min10[,2]), mean))


data_kmedoids10 <- merge(data_kmedoids_ag10,data_keybag,by.x = "Group.1", by.y = "V2",all.x =TRUE)

data_hclust_eucl10 <- merge(data_hclust_eucl_ag10,data_keybag,by.x = "Group.1", by.y = "V2",all.x =TRUE)

data_hclust_min10 <- merge(data_hclust_min_ag10,data_keybag,by.x = "Group.1", by.y = "V2",all.x =TRUE)

```







```{r}
x_medoids10<- as.matrix(data_kmedoids10[,3:12])
y_medoids10<- data_kmedoids10$V1.y

modelprakmedoids10 <- cv.glmnet(x_medoids10, y_medoids10, alpha = 1,nfolds=7,type.measure='auc',family="binomial")

summary<-rbind(summary,list("kmedoids",max(modelprakmedoids10$cvm),10))
#max(modelprakmedoids10$cvm)

x_hclust_eucl10<-as.matrix(data_hclust_eucl10[,3:12])
y_hclust_eucl10<- data_hclust_eucl10$V1.y

modelprahclust_eucl10 <- cv.glmnet(x_hclust_eucl10, y_hclust_eucl10, alpha = 1,nfolds=7,type.measure='auc',family="binomial")

summary<-rbind(summary,list("hclust_euclidean",max(modelprahclust_eucl10$cvm),10))
#max(modelprahclust_eucl10$cvm)


x_hclust_minl10<-as.matrix(data_hclust_min10[,3:12])
y_hclust_minl10<- data_hclust_min10$V1.y

modelprahclust_minl10 <- cv.glmnet(x_hclust_minl10, y_hclust_minl10, alpha = 1,nfolds=9,type.measure='auc',family="binomial")

summary<-rbind(summary,list("hclust_minkowski",max(modelprahclust_minl10$cvm),10))

#max(modelprahclust_minl10$cvm)
print(as.matrix(summary))

```


 In 5 cluster, The hierarchical clustering model using minkowski distance provide best auc value in 7-fold cross validated Lasso . Besides, when the 10 cluster was selected, all models gives good classifier and value converge on 1






