---
title: "Homework2"
author: "Ramazan Yarar"
date: "10/21/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data import

```{r message = FALSE,warning = FALSE }
library (data.table)
library(ggbiplot)
dt_odds<- readRDS("~/Documents/Dersler/Fall-2018/IE-582/df9b1196-e3cf-4cc7-9159-f236fe738215_odd_details.rds")
dt_matches <- readRDS("~/Documents/Dersler/Fall-2018/IE-582/df9b1196-e3cf-4cc7-9159-f236fe738215_matches.rds")
```


# Data manipulation

* Odds table
```{r message = FALSE,warning = FALSE }
# Filtering 5 bookmarkers and final odds

dt_odds<-dt_odds[bookmaker %in% c("10Bet","bet365","Betsafe","bwin","bet-at-home")]

dt_oddsf<-dt_odds[,rnk:=rank(-date,ties.method= "max"),
                            by = list(matchId,betType,oddtype,bookmaker)][rnk==1]

dt_oddss<-dt_oddsf[,.(matchId,bookmaker,betType,oddtype,odd)][betType %in% c("1x2","bts","ou","ha")]

#transpoze for bet
dt_oddst<-dcast(dt_oddss,matchId+bookmaker~betType+oddtype)

#remove null odd (bcs we have enough instance)
setnames(dt_oddst,c('1x2_odd1','1x2_odd2','1x2_oddX'),c("Home","Away","Tie"))
dt_oddst<-na.omit(dt_oddst)
```

* Matches table

```{r message = FALSE,warning = FALSE }
# split score to compare
dt_matches[,c("score1","score2"):= tstrsplit(score,":",fixed=TRUE)]
dt_matches <- dt_matches[,score1:=as.integer(score1)] # to be ensure for calculation
dt_matches <- dt_matches[,score2:=as.integer(score2)]
dt_matches<-na.omit(dt_matches)
# add column to decide Overstatus and Outcometype
dt_matches$totalgoal <- dt_matches$score1 + dt_matches$score2
dt_matches$Overstatus <- ifelse(dt_matches$totalgoal>2.5,"Over", "Under")
dt_matches$Outcomes<-ifelse(dt_matches$score1>dt_matches$score2,"Home",ifelse(dt_matches$score1<dt_matches$score2,"Away","Tie"))
```

* merge table

```{r message = FALSE,warning = FALSE }
# I create setkey to join tables

setkey(dt_oddst,matchId)
setkey(dt_matches,matchId)

# I joined tables 
dt_merged <- merge(dt_oddst,unique(dt_matches),all.x=TRUE)[,.(matchId,Home,Away,Tie,bts_NO,bts_YES,ou_over,ou_under,Overstatus,bookmaker,ha_1,ha_2)]
dt_merged<-na.omit(dt_merged)

dt_merged2 <- merge(dt_oddst,unique(dt_matches),all.x=TRUE)[,.(matchId,Home,Away,Tie,bts_NO,bts_YES,ou_over,ou_under,Outcomes,bookmaker,ha_1,ha_2)]
dt_merged2<-na.omit(dt_merged2)
```




## Task 1

* PCA Analysis
```{r message = FALSE,warning = FALSE }

#filter data (I used the first 250 instance to plot)

dt_merged<-dt_merged[1:250,.(Home,Away,Tie,bts_NO,bts_YES,ou_over,ou_under,ha_1,ha_2,Overstatus,bookmaker)]
dt_merged.pca <- prcomp(dt_merged[,1:9],center = TRUE,scale. = TRUE)
summary(dt_merged.pca)
```


PC1 encompass %43 of total variance, together with PC2 involves %65 of variance. They can provide  insight to data



```{r message = FALSE,warning = FALSE }
ggbiplot(dt_merged.pca,groups=dt_merged$Overstatus,scale = 0,choices = 1:2,obs.scale = 1, var.scale = 1,ellipse = TRUE)
```

 Away , ha_2 and ou_under Tie feature are  located close to each other. This indicates that  variables are correlated with each other. Also Home , ha_1 , ou_over variables are located closely, they are also correlated with each other. These feature may combine to feature reduction
 
 Over and Under results correlated, they generally ploted same regions
 




* MDS plotting for euclidean distance

```{r message = FALSE,warning = FALSE }
d_mcd<-dist(dt_merged[,-10],method = "euclidean")
MDS = cmdscale(d_mcd)
plot(MDS[, 1], MDS[, 2], pch = 21, bg = c("red", "green"))
     #[unclass(dt_merged$Overstatus)])
```

* MDS plotting for manhattan distance

```{r message = FALSE,warning = FALSE }
d_mcd<-dist(dt_merged[,-10],method = "manhattan")
MDS = cmdscale(d_mcd)
plot(MDS[, 1], MDS[, 2], pch = 21, bg = c("red", "green"))
```


  Variables in Manhattan distance has smaller variance than euclidean distance. MDS plotting show us Over and Under have high correlation like PCA results


# Task2

 * PCA Analysis for matches outcomes

```{r message = FALSE,warning = FALSE }
# merge and filter matches outcome
dt_merged2<-dt_merged2[1:250,.(Home,Away,Tie,bts_NO,bts_YES,ou_over,ou_under,ha_1,ha_2,Outcomes,bookmaker)]
dt_merged2.pca <- prcomp(dt_merged2[,1:9],center = TRUE,scale. = TRUE)
#names(dt_bet365.pca)
print(dt_merged2.pca)
summary(dt_merged2.pca)

```


PC1 encompass %43 of total variance, together with PC2 involves %65 of variance. 


```{r message = FALSE,warning = FALSE }

ggbiplot(dt_merged2.pca,groups=dt_merged2$Outcomes,scale = 0,choices = 1:2,obs.scale = 1, var.scale = 1,ellipse = TRUE)


```

Matches results are correlated each other. They are located similarly

## Task 3

* Read Image and Plot

```{r message = FALSE,warning = FALSE }

library(jpeg)
image<-readJPEG("~/Documents/Dersler/Fall-2018/IE-582/image.jpg")
plot(c(0, 512), c(0, 512), type = "n", xlab = "", ylab = "")
rasterImage(image, 0, 0, 512, 512)

```

Data stored as numeric and dimensions are  [1:512, 1:512, 1:3]

* Plot according to color channel

```{r message = FALSE,warning = FALSE }
# break into channel
r <- image[,,1]
g <- image[,,2]
b <- image[,,3]

#Multi plotting
par(mfrow=c(1,3))
image(t(r)[ncol(r):1,nrow(r):1])
image(t(g)[ncol(g):1,nrow(g):1])
image(t(b)[ncol(b):1,nrow(b):1])
```


* Crating Noisy Image

```{r message = FALSE,warning = FALSE }

# add noisy for every channel and pixel
for (i in 1:512)
  {for (j in 1:512)
      {for (z in 1:3)
             {
        image[i,j,z]=image[i,j,z]+runif(1,min=0,max=0.1)
        if(image[i,j,z]>1){
              
                image[i,j,z]=0.99
                     }
                else {image[i,j,z]}
        
                }
        
        }
  
}

# Plot
plot(c(0, 512), c(0, 512), type = "n", xlab = "", ylab = "")
rasterImage(image, 0, 0, 512, 512)
```

* Noisy Image Channel Plotting

```{r message = FALSE,warning = FALSE }
#break into channel
r <- image[,,1]
g <- image[,,2]
b <- image[,,3]

par(mfrow=c(1,3))
image(t(r)[ncol(r):1,nrow(r):1])
image(t(g)[ncol(g):1,nrow(g):1])
image(t(b)[ncol(b):1,nrow(b):1])
```


* PCA analysis for grayscale image

```{r message = FALSE,warning = FALSE }
library(jpeg)
imagea=readJPEG("~/Documents/Dersler/Fall-2018/IE-582/imagegg.jpg")
plot(c(0, 512), c(0, 512), type = "n", xlab = "", ylab = "")
rasterImage(imagea, 0, 0, 512, 512)
```

* Patch Extractions

```{r message = FALSE,warning = FALSE }

# Loop is for extraction
n=0
m=0
mt<-matrix(0,nrow=260100,ncol = 9)
for (i in 2:510)
  {for (j in 2:510)
    { n=0
    m=m+1
      for (k in -1:1) 
         {for (l in -1:1) 
         {
            n=n+1
             mt[m,n]= imagea[i+k,j+l]
       # print(c(i+k,j+l))
       # print(c(m,n))
          
        }  
        
      }
  }
}
 
# PCA analysis
mt.pca<-prcomp(mt[,1:9],center = FALSE,scale. = FALSE)     
summary(mt.pca)
```

PC1 consist  0.99 of total variance. It means that ??t provide insight for %99 of data


* Reconstructing for first component

```{r message = FALSE,warning = FALSE }

#recall loops

deneme<-mt.pca$x[,1]%*%t(mt.pca$rotation[,1])
recall<-matrix(0,nrow=512,ncol = 512)

n=10
m=260101
for (i in 510:2)
  {for (j in 510:2)
    { n=10
     m=m-1
      for (k in 1:-1) 
         {for (l in 1:-1) 
         {
            n=n-1
            if(deneme[m,n]>1)
            {deneme[m,n]=1}
            
            recall[i+k,j+l]=deneme[m,n] 
  
        }  
        
      }
  }
}

#Plot
plot(c(0, 512), c(0, 512), type = "n", xlab = "", ylab = "")
rasterImage(recall, 0, 0, 512, 512)
```

* For second component

```{r message = FALSE,warning = FALSE }

deneme2<-mt.pca$x[,2]%*%t(mt.pca$rotation[,2])

recall2<-matrix(0,nrow=512,ncol = 512)

n=10
m=260101
#mt<-matrix(0,nrow=260100,ncol = 9)
for (i in 510:2)
  {for (j in 510:2)
    { n=10
     m=m-1
      for (k in 1:-1) 
         {for (l in 1:-1) 
         {
            n=n-1
            if(deneme2[m,n]<0)
            {deneme2[m,n]=0}
            
            recall2[i+k,j+l]=deneme2[m,n] 
                  
        }  
        
      }
  }
}


plot(c(0, 512), c(0, 512), type = "n", xlab = "", ylab = "")
rasterImage(recall2, 0, 0, 512, 512)

```

* For third component
```{r message = FALSE,warning = FALSE }
deneme3<-mt.pca$x[,3]%*%t(mt.pca$rotation[,3])

recall3<-matrix(0,nrow=512,ncol = 512)

n=10
m=260101
#mt<-matrix(0,nrow=260100,ncol = 9)
for (i in 510:2)
  {for (j in 510:2)
    { n=10
     m=m-1
      for (k in 1:-1) 
         {for (l in 1:-1) 
         {
            n=n-1
            
            if(deneme3[m,n]<0)
            {deneme3[m,n]=0}
            
            recall3[i+k,j+l]=deneme3[m,n] 
                  
  
          
        }  
        
      }
  }
}


plot(c(0, 512), c(0, 512), type = "n", xlab = "", ylab = "")
rasterImage(recall3, 0, 0, 512, 512)

```

* Task 4c for first component


```{r message = FALSE,warning = FALSE }
#Find Eigenvectors and transform image
tsk4<-mt.pca$rotation[,1]
imagev<-matrix(0,nrow=3,ncol = 3)

m=9
for (i in 1)
  {for (j in 1)
    { 
     
      imagev[i,j]=tsk4[m]*-1
  
    
        }  
        
      }



plot(c(0, 3), c(0, 3), type = "n", xlab = "", ylab = "")
rasterImage(imagev, 0, 0, 3, 3)

```

* Task 4c for second component

```{r message = FALSE,warning = FALSE }


tsk42<-mt.pca$rotation[2,]
imagev2<-matrix(0,nrow=3,ncol = 3)

m=9
#mt<-matrix(0,nrow=260100,ncol = 9)
for (i in 1)
  {for (j in 1)
    { 
     
      imagev2[i,j]=tsk42[m]
  
    
        }  
        
      }



plot(c(0, 3), c(0, 3), type = "n", xlab = "", ylab = "")
rasterImage(imagev2, 0, 0, 3, 3)

```
```{r message = FALSE,warning = FALSE }

tsk43<-mt.pca$rotation[3,]
imagev3<-matrix(0,nrow=3,ncol = 3)

m=9
for (i in 1)
  {for (j in 1)
    { 
     
      imagev3[i,j]=tsk43[m]* -1
  
    
        }  
        
      }



plot(c(0, 3), c(0, 3), type = "n", xlab = "", ylab = "")
rasterImage(imagev3, 0, 0, 3, 3)

```

